---
title: "simpleRegress"
author: "Sophia Castellon"
date: "Fall 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted rsquared for LENGTH~NECK is 0.7476

## Model LENGTH ~ AGE  this is an example

```{r}
simp <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted rsquared for LENGTH~AGE is 0.5073

## Model LENGTH ~ HEADLEN  this is an example

```{r}
simp <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted rsquared for LENGTH~HEADLEN is 0.8434

## Model LENGTH ~ HEADWTH  this is an example

```{r}
simp <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted rsquared for LENGTH~HEADWTH is 0.5321

## Model LENGTH ~ CHEST this is an example

```{r}
simp <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted rsquared for LENGTH~CHEST is 0.7869

## Model LENGTH ~ WEIGHT this is an example

```{r}
simp <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simp)
summary.lm(simp)
```

The adjusted rsquared for LENGTH~WEIGHT is 0.7423

## Compare all models that predict LENGTH other than NECK. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better. 
From worst to best is LENGTH~AGE is 0.5073, LENGTH~HEADWTH is 0.5321, LENGTH~WEIGHT is 0.7423, LENGTH~CHEST is 0.7869, LENGTH~HEADLEN is 0.8434
